---
title: "Day 6 Lecture"
output: html_document
---

# Outline
  - Review Homework 5
  - Recap "data wrangling"
  - More graphics

# Homework 5 Review

# Data munging recap

- read data in carefully
- inspect values
  - value types
  - ranges (of numerics)
  - levels (of factors)
- transform/format variables
  - transform numbers
  - change factor level labels
- reshape
  - melt/gather into long format
    - melt is from the reshape2 package
    - gather is from the tidyr package
  - dcast/spread into wide format
    - dcast is from reshape2
    - spread is from tidyr
- aggregate
  - collapse values across rows, by some other variable/factor
  - dplyr way: group_by() %>% summarize()
  - reshape2 way: dcast(data, cols ~ ., fun.aggregate = some.function)

```{r}
library(dplyr)
library(ggplot2)
library(reshape2)
library(tidyr)

myiris <- iris
myiris$plant <- 1:nrow(myiris)

# get into long format
myiris.long1 <- melt(myiris)
myiris.long1 <- melt(myiris, id.vars = "Species")
myiris.long1 <- melt(myiris, measure.vars = 1:4)

myiris.long1 <- melt(myiris, measure.vars = c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width"), value.name = "cm")

head(myiris.long1)

myiris.long2 <- gather(myiris, variable, cm, Sepal.Length:Petal.Width)
head(myiris.long2)

myiris.long.clean <- separate(myiris.long2, variable, c("flower.part", "dimension"), sep = "\\.")
head(myiris.long.clean)
summary(myiris.long.clean)
?strsplit

# rename factor levels
myiris.tidy <- myiris.long.clean
myiris.tidy$dimension <- as.factor(myiris.tidy$dimension)

levels(myiris.tidy$dimension) <- c("L", "W")
summary(myiris.tidy)

mysleep <- sleep

codes <- c("Bob", "Susan", "Jimmy", "Cora", "Daphne", "Ronan", "Scott", "David Bowie", "Ziggy", "Major Tom")

IDs <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10")

code.book <- data.frame(old = IDs, new = codes, stringsAsFactors = FALSE)

mysleep.subset <- filter(mysleep, !ID %in% c("3", "6")) %>% droplevels()
summary(mysleep.subset)

summary(code.book)

mysleep.subset$ID <- as.character(mysleep.subset$ID)

colnames(code.book) <- c("ID", "newID")

mysleep.subset <- merge(mysleep.subset, code.book)

xtabs(~ ID + newID, mysleep.subset)
```

# More graphics

## faceting
```{r}
myff <- french_fries

myff.long <- gather(myff, taste, rating, potato:painty)

baseplot <- ggplot(myff.long, aes(x = rating))

baseplot + geom_histogram()

baseplot + geom_histogram(aes(fill = taste))

baseplot + geom_histogram() + facet_wrap(~ taste)

head(myiris.tidy)

ggplot(myiris.tidy, aes(x = cm)) + geom_histogram() + facet_wrap(~ Species)

ggplot(myiris.tidy, aes(x = cm)) + geom_histogram() + facet_grid(Species ~ .)

myiris.wide <- dcast(myiris.tidy, plant + Species + dimension ~ flower.part, value.var = "cm")

ggplot(myiris.wide, aes(Sepal, Petal)) + geom_point() + geom_smooth(method = "lm") + facet_grid(Species ~ dimension)

```



```{r}
myanscombe <- anscombe
myanscombe$obs <- 1:nrow(myanscombe)

ans.melt <- melt(myanscombe, measure.vars = 1:8)

ans.melt$newvar <- gsub("(\\d)", "_\\1", ans.melt$variable)

ans.melt <- separate(ans.melt, newvar, c("dimension", "plot"), sep = "_")

ans.cast <- dcast(ans.melt, obs + plot ~ dimension, value.var = "value")

ggplot(ans.cast, aes(x, y)) + geom_point() + geom_smooth(method = "lm", se = FALSE, linetype = 3) + facet_wrap(~ plot)

```

## geom_text
```{r}
ggplot(ans.cast, aes(x, y)) + geom_text(aes(label = obs)) +  facet_wrap(~ plot) #+ geom_point()

```


## geom_bar
```{r}
mysleep.means <- dcast(mysleep, group ~ ., value.var = "extra", fun.aggregate = mean)

mysleep.means2 <- group_by(mysleep, group) %>% summarize(meanExtra = mean(extra))

ggplot(mysleep.means2, aes(1, meanExtra)) + geom_bar(stat = "identity", aes(fill = group), position = "dodge") 


```

## dodging

## bars vs. dots

## errorbars
```{r}

mysleep.means2 <- group_by(mysleep, group) %>% summarize(meanExtra = mean(extra), sdExtra = sd(extra), N = length(extra))

ggplot(mysleep.means2, aes(group, meanExtra)) + geom_bar(stat = "identity", aes(fill = group)) + geom_errorbar(aes(ymax = meanExtra + 1.96*(sdExtra/sqrt(N)), ymin = meanExtra - 1.96*(sdExtra/sqrt(N))))

```

## overplotting

## polishing appearance
