---
title: "Day 4 Lecture"
output: html_document
---

# Today's agenda: lots!

  - Notebooks
  - Merging data
  - Manipulating and transforming variables
  - Reshaping data & "tidy" data
  - Aggregation (e.g., cells means)
  - "Tidy" data
  - Complex objects

# Notebooks vs. scripts
Sweave: the original "reproducible document", mix of LaTeX and R

R Markdown: one of the "cutting edge" formats, especially well-integrated into RStudio, that allows mixing of R code with text, document formatting, links, graphics, etc.

The following is a code chunk, and when `knitr()` is run on this document, it will display both the code and the output.

```{r}
x <- rnorm(10000)
hist(x)  # this is a histogram

t.test(x)
```

You can also insert "inline" results, like the following:

The mean of `x` is `r mean(x)`.

Clicking on the little question mark in the RStudio GUI above this document will point you towards lots of helpful documentation on the details of Markdown and R Markdown documents.

# Data manipulation

## Combining data frame objects
### `rbind()` and `cbind()`
  - rbind
    - glues rows on top of each other
    - expects same columns
    - new and improved `rbind_all()` and `rbind_list()` in `dplyr` package
  - cbind
    - glues columns next to each other
    - need to be the same length    
```{r}
sleep1 <- sleep[sleep$group == 1,  ]
sleep2 <- sleep[sleep$group == 2,  ]
sleep1
sleep2
sleep.reassembled <- rbind(sleep1, sleep2)
sleep.reassembled

# rbind_list
library(dplyr)
sleep2b <- sleep2
sleep2b$new.column <- "other data"
# rbind(sleep1, sleep2b) # doesn't work
data.frame(rbind_list(sleep1, sleep2b)) # from dplyr package
rbind_all(list(sleep1, sleep2b))

sleep.left <- sleep[, 1:2]
sleep.left
sleep.ID <- sleep[, 3]
sleep.ID
sleep.reassembled.2 <- cbind(sleep.left, sleep.ID)
sleep.reassembled.2
```

### `merge()`
```{r}
subject.info <- data.frame(ID = 3:12, gender = c("male", "male", "female", "female", "female", "female", "male", "female", "male", "male"))
subject.info

sleep.merged <- merge(sleep, subject.info)
sleep
sleep.merged
nrow(sleep)
nrow(sleep.merged)

mysleep <- sleep
mysleep$ID <- as.numeric(mysleep$ID)

sleep.merged2 <- merge(mysleep, subject.info, all = TRUE)
sleep.merged2

```
    - simplest: have ONE column in common, named the same thing (with the same values)
    - can get MUCH more complex
    - `nrow()` good check on what was gained/lost during a merge
    
## Creating new columns

```{r}
mysleep
mysleep$Study <- "Study1"
mysleep$noise <- rnorm(20)
mysleep
```

## Simple column transformations

```{r}
mysleep <- sleep
mysleep$extra2 <- mysleep$extra * 2
mysleep$extra3 <- log(mysleep$extra + 10)
mysleep
mysleep$extra2 <- mysleep$extra2 - 4
mysleep
mysleep$extra2 <- NULL  # deletes the column
mysleep
mysleep$extra3 <- NA
mysleep
mysleep$three <- 3
mysleep
# mysleep$four <- c(1, 2, 3) # doesn't work
mysleep$five <- c(1, 2, 3, 4) # does work (recycling)
mysleep

rep(2, 10)
rep(2:4, 10)
rep(2:4, each = 10)

rep(1:2, each = 5)

```

## Dealing with factors (recap from Day 3)
  - levels
  - droplevels
  - recoding
  - reordering
  - as.numeric vs. as.character
  - stringsAsFactors = argument to `read.table()`

### droplevels
```{r}
mysleep <- sleep
summary(mysleep)
levels(mysleep$ID)
summary(mysleep$ID)

mysleep.1 <- subset(mysleep, group == "1")
# mysleep.1 <- mysleep[mysleep$group == "1", ]
mysleep.1 <- subset(mysleep.1, !ID == "3")
# mysleep.1 <- subset(mysleep, group == 1 & !ID == 3)
# mysleep.1 <- mysleep[mysleep$group == 1 & !mysleep$ID == 3, ]
summary(mysleep.1$ID)
summary(mysleep.1$group)

mysleep.2 <- droplevels(mysleep.1)
summary(mysleep.2$ID)
summary(mysleep.2$group)
```

### recoding & reordering
  - relevel, reorder, factor, levels
```{r}
mysleep <- sleep
summary(mysleep)

levels(mysleep$group)
levels(mysleep$group) <- c("pre", "post")  # changes LABELS of the levels
levels(mysleep$group)
mysleep

# levels(mysleep$group) <- c("post", "pre")  # DOESN'T CHANGE ORDER!
# levels(mysleep$group)
# mysleep
# mysleep$group <- factor(mysleep$group, levels = c("post", "pre"))

mysleep$group <- relevel(mysleep$group, "post")  # changes order
levels(mysleep$group)
summary(mysleep)
mysleep

# avoid reorder() except for special cases

mysleep$ID <- factor(mysleep$ID, levels = c("2", "3", "1", "5", "10", "7", "9", "8", "4", "6"))
levels(mysleep$ID)
mysleep

mysleep$IDlevelnum <- as.numeric(mysleep$ID)
mysleep

```

 - `reorder()`
  - takes two vectors, a factor, and an "ordering" vector
  - the "ordering" vector needs to be:
    - the same length
    - numeric values corrsponding to the order of levels in the factor

## Reshaping data
Use the `reshape2` package!!!

```{r}
# install.packages("reshape2")
library(reshape2)
```

### "Long" vs. "wide" data
```{r}
mysleep <- sleep
mysleep
levels(mysleep$group) <- c("pre", "post")
mysleep.cast <- dcast(mysleep, ID ~ group, value.var = "extra")
mysleep.cast

myff <- french_fries
head(myff)
# ?melt.data.frame
myff.melt <- melt(myff, id.vars = 1:4)
myff.melt <- melt(myff, id.vars = c("time", "treatment", "subject", "rep"))
myff.melt <- melt(myff, measure.vars = 5:9)
myff.melt <- melt(myff, measure.vars = c("buttery", "potato", "grassy", "rancid", "painty"))
head(myff.melt)
summary(myff.melt)

myff.melt <- melt(myff, id.vars = 1:4, value.name = "rating", variable.name = "taste")
head(myff.melt)
summary(myff.melt)

recastff <- dcast(myff.melt, time + treatment + subject + rep ~ taste, 
                  value.var = "rating")
summary(recastff)
```
