---
title: "Day 4 Lecture"
output: html_document
---

# Today's agenda: lots!

  - Notebooks
  - Merging data
  - Manipulating and transforming variables
  - Reshaping data & "tidy" data
  - Aggregation (e.g., cells means)
  - "Tidy" data
  - Complex objects

# Notebooks vs. scripts
Sweave: the original "reproducible document", mix of LaTeX and R

R Markdown: one of the "cutting edge" formats, especially well-integrated into RStudio, that allows mixing of R code with text, document formatting, links, graphics, etc.

The following is a code chunk, and when `knitr()` is run on this document, it will display both the code and the output.

```{r}
x <- rnorm(10000)
hist(x)  # this is a histogram

t.test(x)
```

You can also insert "inline" results, like the following:

The mean of `x` is `r mean(x)`.

Clicking on the little question mark in the RStudio GUI above this document will point you towards lots of helpful documentation on the details of Markdown and R Markdown documents.

# Data manipulation

## Combining data frame objects
### `rbind()` and `cbind()`
- rbind
    - glues rows on top of each other
    - expects same columns
    - new and improved `rbind_all()` and `rbind_list()` in `dplyr` package
  - cbind
    - glues columns next to each other
    - need to be the same length    
```{r}
sleep1 <- sleep[sleep$group == 1,  ]
sleep2 <- sleep[sleep$group == 2,  ]
sleep1
sleep2
sleep.reassembled <- rbind(sleep1, sleep2)
sleep.reassembled

# rbind_all
library(dplyr)
sleep2b <- sleep2
sleep2b$new.column <- "other data"
# rbind(sleep1, sleep2b) # doesn't work
rbind_list(sleep1, sleep2b) # from dplyr package

sleep.left <- sleep[, 1:2]
sleep.left
sleep.ID <- sleep[, 3]
sleep.ID
sleep.reassembled.2 <- cbind(sleep.left, sleep.ID)
sleep.reassembled.2
```

### `merge()`
```{r}
subject.info <- data.frame(ID = 3:12, gender = c("male", "male", "female", "female", "female", "female", "male", "female", "male", "male"))
subject.info

sleep.merged <- merge(sleep, subject.info)
sleep
sleep.merged
nrow(sleep)
nrow(sleep.merged)

mysleep <- sleep
mysleep$ID <- as.numeric(mysleep$ID)

sleep.merged2 <- merge(mysleep, subject.info, all = TRUE)
sleep.merged2

```
    - simplest: have ONE column in common, named the same thing (with the same values)
    - can get MUCH more complex
    - `nrow()` good check on what was gained/lost during a merge
    
## Creating new columns

```{r}
mysleep
mysleep$Study <- "Study1"
mysleep$noise <- rnorm(20)
mysleep
```

## Simple column transformations

```{r}
mysleep <- sleep
mysleep$extra2 <- mysleep$extra * 2
mysleep$extra3 <- log(mysleep$extra + 10)
mysleep
mysleep$extra2 <- mysleep$extra2 - 4
mysleep
mysleep$extra2 <- NULL  # deletes the column
mysleep
mysleep$extra3 <- NA
mysleep
mysleep$three <- 3
mysleep
# mysleep$four <- c(1, 2, 3) # doesn't work
mysleep$five <- c(1, 2, 3, 4) # does work (recycling)
mysleep
```

## Dealing with factors (recap from Day 3)
  - levels
  - droplevels
  - recoding
  - reordering
  - as.numeric vs. as.character
  - stringsAsFactors = argument to `read.table()`

### droplevels
```{r}
mysleep <- sleep
summary(mysleep)
levels(mysleep$ID)
summary(mysleep$ID)

mysleep.1 <- subset(mysleep, group == "1")
mysleep.1 <- subset(mysleep.1, !ID == "3")
# mysleep.1 <- subset(mysleep, group == 1 & !ID == 3)
# mysleep.1 <- mysleep[mysleep$group == 1 & !mysleep$ID == 3, ]
summary(mysleep.1$ID)
summary(mysleep.1$group)

mysleep.2 <- droplevels(mysleep.1)
summary(mysleep.2$ID)
summary(mysleep.2$group)
```

### recoding & reordering
  - relevel, reorder, factor, levels
```{r}
mysleep <- sleep
summary(mysleep)

levels(mysleep$group)
levels(mysleep$group) <- c("pre", "post")  # changes LABELS of the levels
levels(mysleep$group)
mysleep

# levels(mysleep$group) <- c("post", "pre")  # DOESN'T CHANGE ORDER!
# levels(mysleep$group)
# mysleep

mysleep$group <- relevel(mysleep$group, "post")  # changes order
levels(mysleep$group)
summary(mysleep)
mysleep

# avoid reorder() except for special cases

mysleep$ID <- factor(mysleep$ID, levels = c("2", "3", "1", "5", "10", "7", "9", "8", "4", "6"))
levels(mysleep$ID)
mysleep

mysleep$IDlevelnum <- as.numeric(mysleep$ID)
mysleep

```

 - `reorder()`
  - takes two vectors, a factor, and an "ordering" vector
  - the "ordering" vector needs to be:
    - the same length
    - numeric values corrsponding to the order of levels in the factor

## Reshaping data
Use the `reshape2` package!!!

```{r}
# install.packages("reshape2")
library(reshape2)
```

### "Long" vs. "wide" data
```{r}
mysleep <- sleep
mysleep
levels(mysleep$group) <- c("pre", "post")
mysleep.cast <- dcast(mysleep, ID ~ group, value.var = "extra")
mysleep.cast

myff <- french_fries
head(myff)
# ?melt.data.frame
myff.melt <- melt(myff, id.vars = 1:4)
myff.melt <- melt(myff, id.vars = c("time", "treatment", "subject", "rep"))
myff.melt <- melt(myff, measure.vars = 5:9)
myff.melt <- melt(myff, measure.vars = c("buttery", "potato", "grassy", "rancid", "painty"))
head(myff.melt)
summary(myff.melt)

myff.melt <- melt(myff, id.vars = 1:4, value.name = "rating", variable.name = "taste")
head(myff.melt)
summary(myff.melt)

recastff <- dcast(myff.melt, time + treatment + subject + rep ~ taste, 
                  value.var = "rating")
summary(recastff)
```

## Aggregation (with the `reshape2` package)

```{r}
aggregateff <- dcast(myff.melt, treatment ~ taste, 
                  value.var = "rating", fun.aggregate = mean, na.rm = TRUE)

aggregateff

aggregateff2 <- dcast(myff.melt, treatment + taste ~ ., 
                  value.var = "rating", fun.aggregate = mean, na.rm = TRUE)
colnames(aggregateff2)[3] <- "mean.rating"

aggregateff2

```

## Aggregation (with the `dplyr` package)

```{r}
aggregateff3 <- group_by(myff.melt, treatment, taste) %>% summarize(mean.rating = mean(rating, na.rm = TRUE))

aggregateff3
identical(aggregateff2, aggregateff3)
identical(aggregateff2, as.data.frame(aggregateff3))

```

## More `dplyr` tips

### Forward-pipe operator (from `magrittr` package)
["This is not a pipe"](https://en.wikipedia.org/wiki/The_Treachery_of_Images)

  - `%>%` operator
  - effect: pass the results of the thing on the left to the first argument of the thing on the right

```{r}
x <- 1:10
mean(log(x + 2))

(x + 2) %>% log() %>% mean()

summarize(group_by(myff.melt, treatment, taste), mean.rating = mean(rating, na.rm = TRUE))

group_by(myff.melt, treatment, taste) %>%
  summarize(mean.rating = mean(rating, na.rm = TRUE))

```

### Subsetting with `dplyr`

```{r}
sleep[sleep$group == 1, ]
filter(sleep, group == 1) # same results

sleep[sleep$group == 1 & sleep$extra > 0.2, ]
filter(sleep, group == 1, extra > 0.2) # same results

sleep[, c("ID", "extra")]
select(sleep, ID, extra)
select(sleep, 3, 1)
select(sleep, c(3, 1))
# select(sleep, c("ID", "extra")) # doesn't work

```

### Alternative package for "tidying" data

    - `tidyr`
    - https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html
    - http://vita.had.co.nz/papers/tidy-data.html

# Data cleaning

## A guideline template for data cleaning
  1. Get the data to read in to R
  2. Fix column names
  3. Check all columns to make sure data is getting read in correctly
  4. Check NAs, decide what to do with them
  5. Remove excessive rows or other "defective" rows
  6. Re-code factors as needed
  
### Reading data
  - find a good `read.X()` function, where X = your type of data
    - `read.table()` is the basic one
    - `read.csv()` and `read.delim()` are convenience "wrappers" for `read.table()`, for CSV and tab-delimited files, respectively
    - `read.spss()` and `read.dta()` from the `foreign` package for SPSS and Stata files
  - for messier or more primitive kinds of files in weird formats, use `readLines()`, or `scan()` to pull in the data in a very raw format, which you will then need to format in R

### Fix column names

```{r}
mysleep <- sleep
oldnames <- colnames(mysleep)
mycolnames <- c("extraSleep", "experimentalGroup", "subjectID")
colnames(mysleep) <- mycolnames
data.frame(oldnames, mycolnames)

```

### Check all columns

### Check NAs
  - use summary()

### Remove rows/columns as needed

### Re-code factors as needed

# Complex objects

  - use `names()` and `str()` to look at the structure of a complex object
  - most complex objects have a list-like structure
  - S3 vs. S4 classes
    - `$` notation works for S3
    - Try `@` instead for S4 classes
    - For more on S4: http://adv-r.had.co.nz/OO-essentials.html

```{r}
myff <- french_fries
results <- lm(potato ~ treatment, data = myff)
results
summary(results)
plot(results)
class(results)
names(results)
str(results)
plot(results$residuals)
results.summary <- summary(results)
names(results.summary)
results.summary$adj.r.squared
results.summary$coefficients
class(results.summary$coefficients)
condition.pvals <- results.summary$coefficients[2:3, "Pr(>|t|)"]
condition.pvals

str(summary(results))
# ?lm
```

The adjusted $R^2$ for this model is `r results.summary$adj.r.squared`.
