---
title: "Day 4 Practice"
output: html_document
---

1. Read in the (modified) database of phoneme pairs from Wedel, Kaplan, and Jackson (2013).  The relative path from the working directory of this project is: "scottdata/CognitionPaperModifiedData.csv"

```{r}
library(reshape2)
mydata <- read.csv("../scottdata/CognitionPaperModifiedData.csv")
```

2. Read in the data using the WRONG function.  What does it look like?  Re-read with the right function for the rest of the practice exercises.

```{r}
mydata <- read.delim("../scottdata/CognitionPaperModifiedData.csv")
head(mydata)
mydata <- read.csv("../scottdata/CognitionPaperModifiedData.csv")
```

3. Change some (or all) of the column names.  Provide a table that links original column names with the current/new column names.

```{r}
mydata2 <- mydata
(old.cols <- colnames(mydata2))
colnames(mydata2)[c(1, 3)] <- c("Language", "Phonemes")
(new.cols <- colnames(mydata2))
(colname.correspondences <- data.frame(old.cols, new.cols))

```

4. Locate NAs.  Remove all rows containing NAs.
   (Hint: try using functions `is.na()`, `complete.cases()`, `no.omit()`, and/or `anyNA()`)

```{r}
summary(mydata)
trouble.column <- mydata$HiSegType
mydata <- mydata[!is.na(trouble.column), ]

mydata.cc <- mydata[complete.cases(mydata), ]
nrow(mydata)
nrow(mydata.cc)
summary(mydata.cc)

missing <- is.na(mydata$AllMinPairs)
mydata.cc2 <- mydata[!missing, ]

mydata.cc3 <- na.omit(mydata)

identical(mydata.cc, mydata.cc2)
identical(mydata.cc2, data.frame(mydata.cc3))

```

5. Recode the "Corpus" variable in the following way:
Am = American
Du = Dutch
Fr = French
Ge = German
HK = Cantonese
Ko = Korean
RP = British
Sl = Slovakian
Sp = Spanish

```{r}
(old.langs <- levels(mydata$Corpus))
(new.langs <- c("American", "Dutch", "French", "German", "Cantonese", "Korean", "British", "Slovakian", "Spanish"))
(lang.table <- data.frame(old.langs, new.langs))
levels(mydata$Corpus) <- new.langs
```

6. Create a "long" version of this data, where `WithinMinPairs` and `BetweenMinPairs` are all in one `MinPairs` variable, and there is a column called `MinPairType` that labels whether they are "within" or "between".

```{r}
mydata.long <- melt(mydata, measure.vars = c("WithinMinPairs", "BetweenMinPairs"), variable.name = "MinPairType", value.name = "MinPairs")
```

7. Log-transform this new `MinPairs` variable.

```{r}
mydata.long$MinPairs <- log(mydata.long$MinPairs)
head(mydata.long)
```

8. Reshape the data back to "wide" format with `WithinMinPairs` and `BetweenMinPairs` (now log-transformed) are in different columns.

```{r}
mydata.wide <- dcast(mydata.long, ... ~ MinPairType, value.var = "MinPairs")
head(mydata.wide)

small <- data.frame(factor = c(rep("red", 10), rep("blue", 10)), numbers = rnorm(20))
small
```

9. Create a new categorical variable that labels "low", "mid", and "high" values of the (log-transformed) `WithinMinPairs` variable.

```{r}
summary(mydata.wide$WithinMinPairs[!is.infinite(mydata.wide$WithinMinPairs)])
summary(mydata.wide$WithinMinPairs[!is.infinite(mydata.wide$WithinMinPairs)])
thresholds <- c(2.639, 5.074)
thresholds <- as.numeric(quantile(mydata.wide$WithinMinPairs[!is.infinite(mydata.wide$WithinMinPairs)], na.rm = TRUE)[c(2, 4)])

mydata.wide$Cat <- "high"
mydata.wide$Cat[mydata.wide$WithinMinPairs <= thresholds[2]] <- "mid"
mydata.wide$Cat[mydata.wide$WithinMinPairs <= thresholds[1]] <- "low"

mydata.wide$Cat2 <- ifelse(mydata.wide$WithinMinPairs > thresholds[2], "high", ifelse(mydata.wide$WithinMinPairs <= thresholds[1], "low", "mid"))

```

10. Reorder the levels of this categorical variable.

```{r}
mydata.wide$Cat <- factor(mydata.wide$Cat)
levels(mydata.wide$Cat)
mydata.wide$Cat <- factor(mydata.wide$Cat, levels = c("low", "mid", "high"))
levels(mydata.wide$Cat)
xtabs(~ Cat + Cat2, mydata.wide)

```

11. Get the mean value of (the log-transformed) `WithinMinPairs` variable for each language in the `Corpus` factor.

12. Run a logistic regression, using `glm()`, where "Merged" status is the (binary) dependent variable, and "WithinMinPairs" is the predictor.
  - Hint #1: the basic usage of `glm()` is:
`
results <- glm(DV ~ predictor1 + predictor2 + ...  ,                data = dataframe.object,
               family = "binomial")
`
  - Hint #2: you need the DV to be 0s and 1s, not a factor.
  
13. Use `summary()` to display the results.

14. Using INLINE code, report the estimate, standard error, and p-value for the WithinMinPairs coefficient.  Hint: save the results as objects first, then put in inline code.

EXTRA CREDIT: figure out how to display the above results with only a few decimal places

HINT: use `str()` to figure out how to extract these values.

