---
title: "Day 9 Lecture"
output: html_document
---

# Outline
  - Review homework and loops
  - Writing functions (basics)
  - Vectorization
  - Writing functions (more)

# Functions
  - making a function
  - anonymous functions
  - environments
  - making functions available
  - debugging

## Function basics
  - basic syntax very simple!

```{r eval = FALSE}
my.function <- function(arg1, arg2, ...) {
  #FUNCTION BODY  
}

```

  - "anonymous" functions = "one-use" functions
  
```{r}
library(reshape2)
library(dplyr)
mysleep <- sleep
levels(mysleep$group) <- c("pre", "post")
mysleep.wide <- dcast(mysleep, ID ~ group, value.var = "extra")
mysleep.wide

double.max <- function(x) {
  output <- max(x) * 2  
  output
}

double.max <- function(x) {
  max(x) * 2  
}

#double.max <- function(x) { max(x) * 2 }
#double.max <- function(x) max(x) * 2 


double.max(1:10)

mysleep.wide$newcol <- apply(mysleep.wide[, 2:3], 1, double.max)
mysleep.wide

# anonymous
mysleep.wide$newcol2 <- apply(mysleep.wide[, 2:3], 1, function(x) { max(x) * 2 })
mysleep.wide

# dplyr
mysleep.wide <- 
  mysleep.wide %>%
  rowwise() %>%
  mutate(newcol3 = double.max(c(pre, post))) # or newcol3 = max(pre, post) * 2
mysleep.wide

# dplyr with melt (good for lots of columns)
mysleep.wide %>%
  melt(measure.vars = 2:3) %>%
  group_by(ID) %>%
  mutate(newcol4 = double.max(value)) %>%
  dcast(ID + newcol + newcol2 + newcol3 + newcol4 ~ variable, value.var = "value")


```


# Vectorization overview
  - sometimes faster (not always)
  - sometimes easier to think about (not always)
  - split-apply-combine

```{r}
x <- 1:10

x * 3

```


# The `apply` family

## `apply`

- good for "row-wise" operations over data frames (or column-wise)
- give it only the columns/rows you want to operate on
- the `MARGIN` argument says whether it's row-wise (=1) or column-wise (=2)
- the `FUN` argument says what to do with each row/column-wise vector

```{r}
minpair <- read.csv("scottdata/CognitionPaperFinalData.csv")
head(minpair)
colnames(minpair)

minpair$allmean <- apply(minpair[, 10:17], MARGIN = 1, FUN = mean)
head(minpair)

```

## other members of the `apply` family
  - generally avoid, if possible!

### `lapply`
  - "list" apply
  - does something to each member of a list/vector
  - returns a list

### `tapply`
  - "table" apply
  - does something to data in "cells" by a combination of factors
  - returns a table

### `sapply`
  - "simple" apply
  - returns a vector
  - usually what you want instead of `lapply`

# A better way: `plyr` and `dplyr`
  - best for data frames
  
## `plyr`
  - the `plyr` package is older, slower, not under very active development
  - instantiates a "split-apply-combine" philosophy
  - handy paper here: http://www.jstatsoft.org/v40/i01
  - series of related functions
    - XYply, where X = "thing you give it" and Y = "thing it gives you back"
    - `ldply()` = you give it a list, it gives you back a data frame
    - `daply()` = you give it a data frame, it gives you back an array
    - etc.

```{r eval = FALSE}
# library(plyr)
my.function <- (chunk)

results <- ddply(mydata, c("factor1", "factor1"), my.function)

```


## `dplyr`
  - the most common usage for `plyr` is `ddply()` -- give it a data frame and get a data frame back
  - the `dplyr` package takes this and runs with it
  - DON'T library both packages in the same session (unless you really need to and know what you're doing)
  - `dplyr` is implemented to be much faster
  - `dplyr` is the future...
  
### `dplyr` usage
  
  - "verbs": `mutate`, `filter`, `select`, `summarise` (or `summarize`), `arrange`
  - use `group_by()` to set the groups over which something (like `summarize` will apply)
  - `mutate` = "transform"
  - `filter` = "get some of the rows"
  - `select` = "get some of the columns"
  - `summarise` = boil things down to some kind of summary
  - `do` = general purpose!
  
```{r}
library(dplyr)

N <- 5
mydataframes <- list()
length(mydataframes) <- N
for(set.num in 1:N) {
  mydataframes[[set.num]] <- read.csv(paste("sampledata", set.num, ".csv", sep = ""))
}

library(dplyr)
mydata <- rbind_all(mydataframes)  

my.ttest.results <- data.frame(mean1 = rep(NA, 5), mean2 = NA, t = NA, df = NA, p = NA)
for(counter in 1:5) {
  this.ttest <- t.test(mydata$X1[mydata$set == counter], mydata$X2[mydata$set == counter])
  this.mean1 <- this.ttest$estimate[1]
  this.mean2 <- this.ttest$estimate[2]
  this.t <- this.ttest$statistic
  this.df <- this.ttest$parameter
  this.pval <- this.ttest$p.value
  my.ttest.results[counter, ] <- c(this.mean1, this.mean2, this.t, this.df, this.pval)
}

get.ttest.results <- function(data, x, y) {
  this.ttest <- t.test(data[[x]], data[[y]])
  this.mean1 <- this.ttest$estimate[1]
  this.mean2 <- this.ttest$estimate[2]
  this.t <- this.ttest$statistic
  this.df <- this.ttest$parameter
  this.pval <- this.ttest$p.value
  my.ttest.results <- data.frame(this.mean1, this.mean2, this.t, this.df, this.pval)
  my.ttest.results
}
get.ttest.results2 <- group_by(mydata, set) %>% do(get.ttest.results(., "X1", "X2"))

```

# More stuff on functions

## Environments
  - R can look "up" through environments, but not "down"
  - the body of a function is an environment, nested under the environment where it is called

```{r}
double.max <- function(x) {
  output <- max(x) * 2  
  print(environment())
  cat("\nThis is an awesome function\n")
  print(output)
  return(NULL)
}

double.max <- function(x) {
  output <- max(x) * 2
  y <- 6
  output
}

```


## Making functions available
  - in the notebook/script
  - `source()`
    - runs all code in another file

```{r}
source("day9/anotherfunction.R")
```

  - package!
    - some package-making tutorials:
      - http://hilaryparker.com/2014/04/29/writing-an-r-package-from-scratch/
      - http://r-pkgs.had.co.nz/
      - https://support.rstudio.com/hc/en-us/articles/200486488-Developing-Packages-with-RStudio
      - http://cran.r-project.org/doc/contrib/Leisch-CreatingPackages.pdf
      - and lots more

## Debugging
  - run as a non-function
  - insert "test" code
  - use `browser()`, `debug()`, and `undebug()`
  
```{r}
my.analysis <- function() {
  library(dplyr)
  if(require(mice)) {
    print("run mice code")
  } else {
    print("run some alternative code")
  }   
}

my.analysis2 <- function() {
  library(dplyr)
  if(library(mice)) {  # throws an error
    print("run mice code")
  } else {
    print("run some alternative code")
  }   
}


mysleep <- sleep

my.pvals.v1 <- function(data) {
  test.results <- t.test(extra ~ group, data = data)
  pval <- test.results$p.value
  if (pval < .05) {
    output <- "p < .05, significant, thank god"
  } else {
    output <- paste("dammit, p is =", round(pval, 3))
  }
  output
}

my.pvals.v1(mysleep)
mysleep2 <- mysleep[-c(3, 17), ]
my.pvals.v1(mysleep2)
mysleep3 <- mysleep
colnames(mysleep3) <- c("time.sleep", "condition", "subject")
my.pvals.v1(mysleep3)

my.pvals.v2 <- function(data, DV, factor) {
  test.results <- t.test(DV ~ factor, data = data)
  pval <- test.results$p.value
  if (pval < .05) {
    output <- "p < .05, significant, thank god"
  } else {
    output <- paste("dammit, p is =", round(pval, 3))
  }
  output
}

my.pvals.v2(mysleep3, DV = "time.sleep", factor = "condition")



```

### "fake function" debugging

```{r}
my.pvals.v2(mysleep3, DV = "time.sleep", factor = "condition")

my.pvals.v2 <- function(data, DV, factor) {
  
  data <- mysleep3
  DV <- "time.sleep"
  factor <- "condition"
  
  test.results <- t.test(DV ~ factor, data = data)
  pval <- test.results$p.value
  if (pval < .05) {
    output <- "p < .05, significant, thank god"
  } else {
    output <- paste("dammit, p is =", round(pval, 3))
  }
  output
  
  
  
}



```

### insert "print" statements


```{r}
my.pvals.v2(mysleep3, DV = "time.sleep", factor = "condition")

my.pvals.v2 <- function(data, DV, factor) {
  print(DV)
  test.results <- t.test(DV ~ factor, data = data)
  pval <- test.results$p.value
  if (pval < .05) {
    output <- "p < .05, significant, thank god"
  } else {
    output <- paste("dammit, p is =", round(pval, 3))
  }
  output
  
  
  
}


```

### with debugging utilities
  - `browser()`
  - `debug()`, `undebug()`

```{r}
my.pvals.v2(mysleep3, DV = "time.sleep", factor = "condition")

my.pvals.v2 <- function(data, DV, factor) {
  #browser()
  test.results <- t.test(DV ~ factor, data = data)
  pval <- test.results$p.value
  if (pval < .05) {
    output <- "p < .05, significant, thank god"
  } else {
    output <- paste("dammit, p is =", round(pval, 3))
  }
  output  
}

debug(my.pvals.v2)
undebug(my.pvals.v2)

my.pvals.v3 <- function(data, my.formula) {
  #browser()
  test.results <- t.test(my.formula, data = data)
  pval <- test.results$p.value
  if (pval < .05) {
    output <- "p < .05, significant, thank god"
  } else {
    output <- paste("dammit, p is =", round(pval, 3))
  }
  output  
}

my.pvals.v3(mysleep3, time.sleep ~ condition)

my.pvals.v4 <- function(data, DV, factor) {
  formula.string <- paste(DV, "~", factor)
  my.formula <- as.formula(formula.string)
  test.results <- t.test(my.formula, data = data)
  pval <- test.results$p.value
  if (pval < .05) {
    output <- "p < .05, significant, thank god"
  } else {
    output <- paste("dammit, p is =", round(pval, 3))
  }
  output  
}

my.pvals.v4(mysleep3, DV = "time.sleep", factor = "condition")


```

### dealing with user error

``` {r}
my.pvals.v4 <- function(data, DV, factor) {
  if(!is.data.frame(data)) {
    stop("Please supply a data frame, dear user") # can use warning() as well, does not throw error
  }
  formula.string <- paste(DV, "~", factor)
  my.formula <- as.formula(formula.string)
  test.results <- t.test(my.formula, data = data)
  pval <- test.results$p.value
  if (pval < .05) {
    output <- "p < .05, significant, thank god"
  } else {
    output <- paste("dammit, p is =", round(pval, 3))
  }
  output  
}

my.pvals.v4(as.matrix(mysleep3), DV = "time.sleep", factor = "condition")


```
