---
title: "Day 5 Lecture"
output: html_document
---

# Outline
  - Review Homework 4
  - Finish "data wrangling"
    - data aggregation/summarization
    - tidy data
    - complex objects
  - Graphics
    - base graphics
    - ggplot2 basics
    - graphics "devices" and saving graphics

# Homework 4 Review


## Aggregation (with the `reshape2` package)

```{r}
library(reshape2)

myff <- french_fries
myff.melt <- melt(myff, id.vars = 1:4, value.name = "rating", variable.name = "taste")
head(myff.melt)

head(dcast(myff.melt, time + treatment + subject + rep ~ taste, value.var = "rating"))
head(dcast(myff.melt, ... ~ taste, value.var = "rating"))

aggregateff <- dcast(myff.melt, taste ~ ., value.var = "rating", fun.aggregate = mean, na.rm = TRUE)

aggregateff

aggregateff <- dcast(myff.melt, treatment ~ taste, value.var = "rating", fun.aggregate = mean, na.rm = TRUE)
aggregateff


aggregateff2 <- dcast(myff.melt, treatment + taste ~ ., 
                  value.var = "rating", fun.aggregate = mean, na.rm = TRUE)
aggregateff2
colnames(aggregateff2)[3] <- "mean.rating"
aggregateff2

# template:
# dcast(yourdata, columns + that + stay + columns ~ cast + columns,
#       value.var = "measure.to.aggregate", 
#       fun.aggregate = function.to.aggregate,
#       agg.function.options...)

```

## Aggregation (with the `dplyr` package)

```{r}
library(dplyr)
head(myff.melt)
aggregateff3 <- group_by(myff.melt, treatment, taste) %>% summarize(mean = mean(rating, na.rm = TRUE), sd = sd(rating, na.rm = TRUE), max.rating = max(rating, na.rm = TRUE), N = length(which(!is.na(rating))))
as.data.frame(aggregateff3)
# add informative comment here

as.data.frame(aggregateff3)
identical(aggregateff2, aggregateff3)
identical(aggregateff2, as.data.frame(aggregateff3))

```

## More `dplyr` tips

### Forward-pipe operator (from `magrittr` package)
["This is not a pipe"](https://en.wikipedia.org/wiki/The_Treachery_of_Images)

  - `%>%` operator
  - effect: pass the results of the thing on the left to the first argument of the thing on the right

```{r}
x <- 1:10
mean(log(x + 2))

(x + 2) %>% log() %>% mean()

summarize(group_by(myff.melt, treatment, taste), mean.rating = mean(rating, na.rm = TRUE))

# same as above, using the %>% operator
agg.results <- group_by(myff.melt, treatment, taste) %>%
  summarize(mean.rating = mean(rating, na.rm = TRUE))

agg.results

```

### Subsetting with `dplyr`

```{r}
sleep
sleep[sleep$group == 1, ]
filter(sleep, group == 1) # same results

sleep[sleep$group == 1 & sleep$extra > 0.2, ]
filter(sleep, group == 1, extra > 0.2) # same results

sleep[ , c("ID", "extra")]
select(sleep, ID, extra)
select(sleep, 3, 1)
select(sleep, c(3, 1))
# select(sleep, c("ID", "extra")) # doesn't work

```

### Alternative package for "tidying" data

    - `tidyr`
    - https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html
    - http://vita.had.co.nz/papers/tidy-data.html
    
```{r}
library(tidyr)
mysleep <- sleep
mysleep
levels(mysleep$group) <- c("pre", "post")

mysleep.cast <- dcast(mysleep, ID ~ group, value.var = "extra")
mysleep.cast

spread(mysleep, group, extra)

melt(mysleep.cast, measure.vars = c("pre", "post"), variable.name = "time", value.name = "extra.sleep")
gather(mysleep.cast, time, extra.sleep, c(pre,post))
```

# Data cleaning

## A guideline template for data cleaning
  1. Get the data to read in to R
  2. Fix column names
  3. Check all columns to make sure data is getting read in correctly
  4. Check NAs, decide what to do with them
  5. Remove excessive rows or other "defective" rows
  6. Re-code factors as needed
  
### Reading data
  - find a good `read.X()` function, where X = your type of data
    - `read.table()` is the basic one
    - `read.csv()` and `read.delim()` are convenience "wrappers" for `read.table()`, for CSV and tab-delimited files, respectively
    - `read.spss()` and `read.dta()` from the `foreign` package for SPSS and Stata files
  - for messier or more primitive kinds of files in weird formats, use `readLines()`, or `scan()` to pull in the data in a very raw format, which you will then need to format in R

### Fix column names

```{r}
mysleep <- sleep
oldnames <- colnames(mysleep)
mycolnames <- c("extraSleep", "experimentalGroup", "subjectID")
colnames(mysleep) <- mycolnames
data.frame(oldnames, mycolnames)

```

### Check all columns

### Check NAs
  - use summary()

### Remove rows/columns as needed

### Re-code factors as needed

# Complex objects

  - use `names()` and `str()` to look at the structure of a complex object
  - most complex objects have a list-like structure
  - S3 vs. S4 classes
    - `$` notation works for S3
    - Try `@` instead for S4 classes
    - For more on S4: http://adv-r.had.co.nz/OO-essentials.html

```{r}
mysleep
# my.ttest <- t.test(extra ~ group, data = mysleep)
# print(my.ttest)
# str(my.ttest)

myff <- french_fries
results <- lm(potato ~ treatment, data = myff)
results
summary(results)
# plot(results)
class(results)
names(results)
str(results)
plot(results$residuals)
results.summary <- summary(results)
names(results.summary)
results.summary$adj.r.squared
results.summary$coefficients
class(results.summary$coefficients)
condition.pvals <- results.summary$coefficients[c("treatment2", "treatment3"), "Pr(>|t|)"]
condition.pvals

# str(summary(results))
# ?lm
```

The adjusted $R^2$ for this model is `r results.summary$adj.r.squared`.

# Graphics overview

http://accidental-art.tumblr.com/

## Two systems

  - "base" graphics
  - "grid" graphics
  - See *R Graphics* by Paul Murrell for a very thorough "deep-dive" into these implementations:  https://www.stat.auckland.ac.nz/~paul/RG2e/

### Base graphics
  - more "primitive"
  - can also be much faster than grid
  - lots of isolated plot functions
  - lots of weird little parameters to worry about
  - some handy quick-plot functions from base graphics:
    - `plot()` : a generic function that will often produce something interesting, can do scatterplots with a formula `y ~ x`
    - `hist()` : histogram of a vector
    - `boxplot()` : boxplot, using a formula `variable ~ grouping.factor`

### Grid graphics
  - better for more "high-level" plots
  - much more complex underlying system (IMHO)
  - more powerful and extendable
  - the basis of extremely popular packages such as `lattice` and `ggplot2`

## Graphics packages

  - many packages will have some kind of graphics functions in them, for visualizing specific kinds of things
  - these will often be in base graphics, "pre-canned", etc.
  - the two best more general-purpose graphics packages are `lattice` and `ggplot2`
  - `ggplot2` is another Hadley Wickham project, and has the most documentation, examples, and is currently most widely used

# Using `ggplot2`

## A "Grammar of Graphics"
  - based on work by Leland Wilkinson, a "theory" of graphic display of data
  - Hadley implemented this framework as his dissertation
  - `ggplot2` is the overhauled version of the original implementation, and it is still under active development
  - the idea is to implement a general system for mapping data variables to *aesthetics* (Wilkinson's term), i.e., visual dimensions (x, y, color, etc.)
  - allows for high-level control of the structure of a plot, to allow for flexible, powerful visualizations
  
## Basic structure and concepts

  - unlike base graphics, the *value* of `ggplot2` objects is a complex object, and `print()`-ing them renders the graph.
  - the value of the object is built up by adding *layers*
  - start with a call to `ggplot()` that specifies the data and some mappings of variables to aesthetics
  - add "geoms" to actually render something
  - add other things for fine control over graphical parameters, etc.

## Graphing the `iris` package

```{r}
library(ggplot2)
myris <- iris
summary(myris)

ggplot(myris, aes(x = Sepal.Length)) + geom_histogram()

ggplot(myris, aes(x = Sepal.Length)) + geom_histogram(binwidth = .5)

ggplot(myris, aes(x = Sepal.Length)) + geom_histogram(binwidth = .01)

mybaseplot <- ggplot(myris, aes(x = Sepal.Length))

mybaseplot + geom_histogram(binwidth = .2)

names(mybaseplot)
str(mybaseplot)

base.density <- mybaseplot + geom_density(adjust = .5, fill = "grey")

mydensity <- geom_density(adjust = .5, fill = "grey")

mybaseplot + mydensity

```

```{r}
colors()
```

```{r}
pdf("bunch of plots.pdf")
ggplot(myris, aes(x = Sepal.Length, y = Sepal.Width)) + geom_point()

ggplot(myris, aes(x = Sepal.Length, y = Sepal.Width)) + geom_point(color = "papayawhip") + theme_classic()

ggplot(myris, aes(x = Sepal.Length, y = Sepal.Width)) + 
  geom_point(aes(color = Species), size = 10) + 
  theme_minimal()

ggplot(myris, aes(x = Sepal.Length, y = Sepal.Width)) + 
  geom_point(aes(color = Species, size = Petal.Width)) + 
  theme_minimal()

ggplot(myris, aes(x = Sepal.Length, y = Sepal.Width)) + 
  geom_point() + 
  geom_smooth()

ggplot(myris, aes(x = Sepal.Length, y = Sepal.Width)) + 
  geom_point() + 
  geom_smooth(method = "lm")

ggplot(myris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) + 
  geom_point(size = 4) + 
  theme_minimal() +
  geom_smooth(method = "lm")
dev.off()

```

### Saving plots

```{r}
pdf("myplot.pdf", width = 10, height = 4)
ggplot(myris, aes(x = Sepal.Length)) + geom_histogram() + 
  facet_grid(Species ~ .)
dev.off()

?Devices
```

### Faceting

```{r}

ggplot(myris, aes(x = Sepal.Length)) + geom_histogram() + 
  facet_wrap(~ Species)

ggplot(myris, aes(x = Sepal.Length)) + geom_histogram() + 
  facet_grid(Species ~ .)

ggplot(myris, aes(x = Sepal.Length)) + geom_histogram() + 
  facet_grid(Species ~ .)

```

# Practice 5

# Review Practice 5

# Admin round-up

## Weekend homework

## Project proposals due!
